---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}

# set working directory and read in train data

data_train = read.csv("ASD_data_train.csv", sep = ",")
View(data_train)

# install libraries 
library(ggplot2)
library(lme4)
library(simr)

```

```{r}

model <- lmer(CHI_MLU ~ ADOS.y * VISIT.x + MOT_MLU + verbalIQ.y + (1 + VISIT.x|SUBJ), data_train, REML = F)
summary(model)

main_effect_model <- lmer(CHI_MLU ~ ADOS.y + VISIT.x + MOT_MLU + verbalIQ.y + (1 + VISIT.x|SUBJ), data_train, REML = F)

power_interaction = powerSim(model,fixed("ADOS.y:VISIT.x"),nsim=200)
power_ADOS = powerSim(main_effect_model,fixed("ADOS.y"),nsim=200)
power_VISIT = powerSim(main_effect_model,fixed("VISIT.x"),nsim=200)
power_MOT = powerSim(model,fixed("MOT_MLU"),nsim=200)
power_verbalIQ = powerSim(model,fixed("verbalIQ.y"),nsim=200)

power_interaction
# power 100%
power_ADOS
# power 7.5%
power_VISIT
# power 100%
power_MOT
# power 100%
power_verbalIQ
# 100%

```
Assesing power in the model (...) with sample size constant at 61 participants we found x y and z to be sufficiently powered, and q, w and c to be insufficiently powered: 
Variable ADOS (diagnostic value) with assumtion of effect size xx has a power of xxx. 
Variable Visit (time) with assumption of effect size yy has a power of yyy. 
Varibale parental MLU with assumption of effect size zz has a power of zzz.
Variable verbal IQ with assumption of effect size qq has a power of qqq.
To include varibale q, w and c which are insufficiently powered one need more participants in the study. 



### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
# the smaller effects I wish to find, the harder it will be to have power enough and the more observations are needed.  
# smaller effects are more conservative - look for smaller changes. 
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}
summary(main_effect_model)
#effect ADOS: -0.0025 -> same (very small already) -> -0.5
#effect VISIT: 0.19 -> 0.15 -> 0.2
#effect MOT_MLU: 0.343 -> 0.3 -> 0.2
#effect verbal IQ: 0.065 -> 0.05 -> 0.2
summary(model)
#effect interaction: -0.018 -> same (very small already) -> -0.2
#effect MOT_MLU: -0.669
#effect verbalIQ.y: -0.558

# main effects 
fixef(main_effect_model)["ADOS.y"] <- -0.0025
fixef(main_effect_model)["VISIT.x"] <- 0.15
fixef(main_effect_model)["MOT_MLU"] <- 0.3
fixef(main_effect_model)["verbalIQ.y"] <- 0.05

#interaction effects
fixef(model)["ADOS.y:VISIT.x"] <- -0.018
fixef(model)["MOT_MLU"] <- -0.669
fixef(model)["verbalIQ.y"] <- -0.558

# power curve 
curve_visit <- powerCurve(main_effect_model,fixed("VISIT.x"),along = "SUBJ",nsim=200)
curve_visit
plot(curve_visit)
# 13 > 15 > 10

curve_ados <- powerCurve(main_effect_model,fixed("ADOS.y"),along = "SUBJ",nsim=200)
curve_ados
plot(curve_ados)
# has barly no effect 

curve_mot <- powerCurve(main_effect_model,fixed("MOT_MLU"),along = "SUBJ",nsim=200)
curve_mot
plot(curve_mot)
# 17 > 25

curve_verbal <- powerCurve(main_effect_model,fixed("verbalIQ.y"),along = "SUBJ",nsim=200)
curve_verbal
plot(curve_verbal)
# 21 > 20

curve_interact <- powerCurve(model,fixed("ADOS.y:VISIT.x"),along = "SUBJ",nsim=200)
curve_interact
plot(curve_interact)
# 15

curve_mot2 <- powerCurve(model,fixed("MOT_MLU"),along = "SUBJ",nsim=200)
curve_mot2
plot(curve_mot2)
# 6

curve_verbal2 <- powerCurve(model,fixed("verbalIQ.y"),along = "SUBJ",nsim=200)
curve_verbal2
plot(curve_verbal2)
# 8

### Riccardo's clumsy function to simulate new participants
### TO DO points are only notes for myself, so not part of the assignment

createNewData <- function (participants,visits,model){
  # participants is the number of subjects
  # visits is the number of visits
  # TO DO: LOOP THROUGH ALL FE ROWS AND AUTOMATICALLY EXTRACT NAMES OF FIXED EFFECTS AND ESTIMATES
  fe <- fixef(model)
  Intercept <- fe[1] #intercept
  bVisit <- fe[2] #visit
  bDiagnosis <- fe[3] #diagnosis
  bVisitDiagnosis <- fe[4] #visit diagnosis interaction
  # TO DO: INTEGRATE STANDARD ERROR?
  
  # TO DO: LOOP THROUGH ALL VC COMPONENTS AND AUTOMATICALLY EXTRACT NAMES OF EFFECTS AND ESTIMATES
  vc<-VarCorr(model) # variance component
  sigmaSubject <- as.numeric(attr(vc[[1]],"stddev")[1]) # random intercept by subject
  sigmaVisit <- as.numeric(attr(vc[[1]],"stddev")[2]) # random slope of visit over subject
  sigmaResiduals <- as.numeric(attr(vc,"sc"))
  sigmaCorrelation <- as.numeric(attr(vc[[1]],"correlation")[2])
  
  # Create an empty dataframe
  d=expand.grid(Visit=1:visits,Child.ID=1:participants)
  # Randomly sample from a binomial (to generate the diagnosis)
  condition <- sample(rep(0:1, participants/2))
  d$Diagnosis<-condition[d$Child.ID]
  d$Diagnosis[is.na(d$Diagnosis)]<-1
  
  ## Define variance covariance matrices:
  Sigma.u<-matrix(c(sigmaSubject^2,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaVisit^2),nrow=2)
  
  ## generate new fake participants (column1=RandomIntercept, column2=RandomSlope)
  u<-mvrnorm(n=participants,
             mu=c(0,0),Sigma=cov(ranef(model)$Child.ID))
  
  ## now generate fake data:
  ### the outcome is extracted from a gaussian with
  ### the solution to the model's equation as mean and
  ### the residual standard deviation as standard deviation 
  d$CHI_MLU <- rnorm(participants*visits,
                     (Intercept+u[,1]) +
                     (bVisit+u[,2])*d$Visit + 
                     bDiagnosis*d$Diagnosis ,sigmaResiduals)  
  
  return(d)
}
```
Should make more participant for ADOS ???
Should set effect size at .05?

### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.
```{r}
# changing effect sizes to find one where 30 participants give a power of 80%


# visit 
fixef(main_effect_model)["VISIT.x"] <- 0.1
curve_visit <- powerCurve(main_effect_model,fixed("VISIT.x"),along = "SUBJ",nsim=20)
curve_visit
plot(curve_visit)

 # ados 
fixef(main_effect_model)["ADOS.y"] <- -0.03
curve_ados <- powerCurve(main_effect_model,fixed("ADOS.y"),along = "SUBJ",nsim=20)
curve_ados
plot(curve_ados)

# mot
fixef(main_effect_model)["MOT_MLU"] <- 0.23
curve_mot <- powerCurve(main_effect_model,fixed("MOT_MLU"),along = "SUBJ",nsim=20)
curve_mot
plot(curve_mot)

# verbal iq
fixef(main_effect_model)["verbalIQ.y"] <- 0.9
curve_verbal <- powerCurve(main_effect_model,fixed("verbalIQ.y"),along = "SUBJ",nsim=20)
curve_verbal
plot(curve_verbal)

# interaction
fixef(model)["ADOS.y:VISIT.x"] <- -0.01
curve_interact <- powerCurve(model,fixed("ADOS.y:VISIT.x"),along = "SUBJ",nsim=20)
curve_interact
plot(curve_interact)

#interaction model mot 
fixef(model)["MOT_MLU"] <- -0.669
curve_mot2 <- powerCurve(model,fixed("MOT_MLU"),along = "SUBJ",nsim=20)
curve_mot2
plot(curve_mot2)

# interaction model mot 
fixef(model)["verbalIQ.y"] <- -0.558
curve_verbal2 <- powerCurve(model,fixed("verbalIQ.y"),along = "SUBJ",nsim=20)
curve_verbal2
plot(curve_verbal2)

```

